Clustering time series based on probability distributions across temporal granularities

### 2021-11-17 main.Rmd
XXX for Fig 9 addressed

Clustering of the $24$ prototypes <!--and the data subset consisting of $353$ customers--> was conducted with all three distances, JS-NQT, JS-RS and WPD, and is summarised in Figures \ref{fig:opt-cluster-tsne-jsd}, \ref{fig:combined-groups-js} and \ref{fig:parcoord-application}. (XXX move the next sentences down to where you talk about the clustering results for JS) In the following plots, the median is shown by a line, and the shaded region shows the area between the  $25^{th}$ and $75^{th}$ percentiles.<!--All customers with the same color represent the same clustered groups. --> Groups are colored differently whenever they represent different groupings. While plotting the clusters, RS-based scaling is applied to each customer, so that no customer dominates the shape of the summarized clusters because of magnitude. The plotting scales are not displayed since we want to emphasize comparable patterns rather than scales. The idea is that a customer in a cluster may have low total energy usage, but their behavior may be quite similar to a customer with high usage with respect to distributional pattern or significance across cyclic granularities.

@@

The $sindex$ is plotted against the number of clusters for clustering based on JS-based (JS-NQT) distances in Figure \ref{fig:opt-cluster-tsne-jsd}. The $sindex$ falls from around $0.56$ when $k=3$ to $0.54$ when $k$ increases. In our study, we use $k=4, 5$. Figure \ref{fig:groups-4and5} depicts the summarized distributions across $4$ and $5$ clusters in (a) and (b) respectively, and assists in characterizing each cluster. Groups sA-1, A-2, and A-3 have profiles that correspond to B-1, B-2, and B-3 respectively. B-2 (id: 4-9 in Figure \ref{fig:hod-ind-group-png}) and B-1 (id: 21-24) have a $hod$ pattern with a typical morning and evening peak, whereas the other groups have no distinct pattern or high variability throughout the day. A-4 is subdivided into B-4 and B-5, each of which has a distinct shape across $moy$ and $wnwd$. B-4 (id: 16-20) is distinguished by higher energy consumption in the first few months of the year and greater variability in weekend usage. B-5 (id: 1-3) has higher consumption in the middle of the year (winter months) and a similar weekday-weekend inter-quartile range. When $k = 4$ is used, these two groups merge to form A-4, which has a moy profile of higher usage in both the beginning and middle of the year, which is not representative of the individuals. It may be worthwhile to compare Figures \ref{fig:opt-cluster-tsne-jsd} and \ref{fig:hod-ind-group-png} to see if the summarized distributions across groups accurately characterized the groupings. If it has, then the majority of the members of the group should have a similar profile.

@@

```{r groups-4and5, fig.cap = "Summary plots for four and five clusters. The distribution of electricity demand combined for all members over $hod$, $moy$, and $wnwd$ is shown. Groups A-1 (customers XXX), A-2 (customers XXX), and A-3 (customers XXX) profiles correspond to Groups B-1, B-2, and B-3, respectively. Cluster A-4 splits into B-4 (customers XXX) and B-5 (customers XXX)to produce the five clusters, whci h better resolves the $moy$ distribution. B-4 is characterized by higher energy consumption in the first few months of the year and more variation in weekend usage. B-5 (id:1-3) is distinguished by higher consumption in the middle of the year (winter months) and similar weekday-weekend inter-quartile range. When $k = 4$ is used, these two groups merge to form A-4, which has a moy profile of higher usage in both the beginning and middle of the year, which is not representative of the individuals in the group.", fig.height=5}

>>>>

Clustering of the $24$ prototypes <!--and the data subset consisting of $353$ customers--> was conducted with all three distances, JS-NQT, JS-RS and WPD, and is summarised in Figures \ref{fig:opt-cluster-tsne-jsd}, \ref{fig:combined-groups-js} and \ref{fig:parcoord-application}. Groups are colored differently whenever they represent different groupings. While plotting the clusters, RS-based scaling is applied to each customer, so that no customer dominates the shape of the summarized clusters because of magnitude. The plotting scales are not displayed since we want to emphasize comparable patterns rather than scales. The idea is that a customer in a cluster may have low total energy usage, but their behavior may be quite similar to a customer with high usage with respect to distributional pattern or significance across cyclic granularities.

@@

The $sindex$ is plotted against the number of clusters for clustering based on JS-based (JS-NQT) distances in Figure \ref{fig:opt-cluster-tsne-jsd}. The $sindex$ falls from around $0.56$ when $k=3$ to $0.54$ when $k$ increases. In our study, we use $k=4, 5$. Figure \ref{fig:groups-4and5} depicts the summarized distributions across $4$ and $5$ clusters in (a) and (b) respectively, and assists in characterizing each cluster. In the following plots, the median is shown by a line, and the shaded region shows the area between the  $25^{th}$ and $75^{th}$ percentiles. Groups A-1, A-2, and A-3 have profiles that correspond to B-1, B-2, and B-3 respectively. B-2 (id: 4-9 in Figure \ref{fig:hod-ind-group-png}) and B-1 (id: 21-24) have a $hod$ pattern with a typical morning and evening peak, whereas the other groups have no distinct pattern or high variability throughout the day. A-4 is subdivided into B-4 and B-5, each of which has a distinct shape across $moy$ and $wnwd$. B-4 (id: 16-20) is distinguished by higher energy consumption in the first few months of the year and greater variability in weekend usage. B-5 (id: 1-3) has higher consumption in the middle of the year (winter months) and a similar weekday-weekend inter-quartile range. When $k = 4$ is used, these two groups merge to form A-4, which has a moy profile of higher usage in both the beginning and middle of the year, which is not representative of the individuals. It may be worthwhile to compare Figures \ref{fig:opt-cluster-tsne-jsd} and \ref{fig:hod-ind-group-png} to see if the summarized distributions across groups accurately characterized the groupings. If it has, then the majority of the members of the group should have a similar profile.

@@

```{r groups-4and5, fig.cap = "Summary plots for four and five clusters. The distribution of electricity demand combined for all members over $hod$, $moy$, and $wnwd$ is shown. Groups A-1 (customers 21-24), A-2 (customers 4-9), and A-3 (customers 10-15) profiles correspond to Groups B-1, B-2, and B-3, respectively. Cluster A-4 splits into B-4 (customers 16-20) and B-5 (customers 1-3) to produce the five clusters, which better resolves the $moy$ distribution. B-4 is characterized by higher energy consumption in the first few months of the year and more variation in weekend usage. B-5 is distinguished by higher consumption in the middle of the year (winter months) and similar weekday-weekend inter-quartile range. When $k = 4$ is used, these two groups merge to form A-4, which has a moy profile of higher usage in both the beginning and middle of the year, which is not representative of the individuals in the group.", fig.height=5}
